Initial Explorations

### 2019-02-22 initial_exploration.Rmd
update

---
title: "Initial Exploration"
author: "Dan Johnson"
date: "2/3/2019"
output: html_document
---

@@

#Preliminary Exploration of TA Comments Dataset
The first question to answer before making deeper comparisons is **are TAs' comments obviously different between sub-categories?** This page describes several different visual comparisons I made of word frequencies before beginning in-depth text analyses.

*  Comparisons of word frequencies in subject sub-categories, using log-odds ratios
    +  Pearson Correlations of Word Frequencies
    +  Mapping Analyses Visually
    +  Calculating and Plotting Log-Odds Pairs
*  Direct comparison of word frequency lists

##Initial Library and Data Import
```{r Libraries_1}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

```{r}
#This block reads in full CSV into a starting dataframe named "base_data".
#Subsequent analyses are performed on extracted subsets of "base_data".
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

@@

##Direct Comparison Using Only Words Appearing in All Subsets of Comments
I think this is less informative, but it does let me see the data more clearly.

Basic strategy is to take all comments from base_data, unnest tokens, filter stopwords, count and sort remaining terms, then write back to a new datatable called 'sortedwords.comments.all'.  Calculate TOTAL words and store. Then calculate fractional frequencies for each word and store using mutate.

I wrote this block so that, once CSV is imported, it does not require any of the code above. This is an early effort, so likely has some poorly developed code that could be refactored.

```{r}
# Tokenize phrases, remove the stop words listed in R's standard reference file.
base_data_tokenized <- base_data %>% 
  unnest_tokens(word,ta.comment)

# This is an ALTERNATIVE METHOD for pulling out stopwords.  
sortedwords.comments.all <- base_data_tokenized %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.all <- sortedwords.comments.all %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
  all.fraction = (n*100)/total_words.all$total
  )
```

```{r}
#Pull SUBSETs of tokenized data based on code.subject: 
base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1_basic", select = (1:28))
base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2_writing", select = (1:28))
base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3_technical", select = (1:28))
base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4_logic", select = (1:28))
#base_data_tokenized.praise <- subset(base_data_tokenized, code.subject == "5_exposition", select = (1:28))
#base_data_tokenized.misconduct <- subset(base_data_tokenized, code.subject == "6_misconduct", select = (1:28))
#base_data_tokenized.narrative <- subset(base_data_tokenized, code.subject == "12. Narrative Comments", select = (1:28))
#base_data_tokenized.sciname <- subset(base_data_tokenized, code.subject == "Scientific Name", select = (1:28))
```

For each subset, filter stopwords, count words and total, then calculate frequencies.

```{r}
sortedwords.basic <- base_data_tokenized.basic %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

@@

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.basic.fraction <- sortedwords.basic %>% as_tibble() %>% mutate(
  basic.fraction = (n*100)/total_words.basic$total
)
```

```{r}
sortedwords.writing <- base_data_tokenized.writing %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.writing <- sortedwords.writing %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.writing.fraction <- sortedwords.writing %>% as_tibble() %>% mutate(
  writing.fraction = (n*100)/total_words.writing$total
)

sortedwords.technical <- base_data_tokenized.technical %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.technical <- sortedwords.technical %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.technical.fraction <- sortedwords.technical %>% as_tibble() %>% mutate(
  technical.fraction = (n*100)/total_words.technical$total
)

sortedwords.logic <- base_data_tokenized.logic %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate the number of words in total in the subset
total_words.logic <- sortedwords.logic %>% 
  summarize(total = sum(n))

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.logic.fraction <- sortedwords.logic %>% as_tibble() %>% mutate(
  logic.fraction = (n*100)/total_words.logic$total
)
```

Next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed.

**WARNING**: if a word is MISSING from **any subset** the data for that word is pruned from the full dataset. I am losing the unique words this way. **Need to figure out how to join without pruning.**

```{r}
total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
total2 <- merge(total,sortedwords.writing.fraction,by="word")
total3 <- merge(total2,sortedwords.logic.fraction,by="word")
#rename lines that were possibly confusing
names(total3)[2] <- "all.count"
names(total3)[4] <- "technical.count"
names(total3)[6] <- "writing.count"
names(total3)[8] <- "logic.count"

@@

The following six graphs compare frequencies for ~770 words that appear in ALL THREE of the sub-categories.

Graphs 1-3 compare frequencies in individual groups against word frequency overall. Generally, the frequency of a term in the whole set of comments is similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup clearly fall below the trendline Graph 3, suggesting they are less frequent in that sub-category than overall.

```{r}
ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")

@@

ggplot(total6, aes(`all.fraction`,`logic.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
```

Graphs 4-6 are pair-wise comparisons of frequencies for sub-categories. Unlike the comparisons to overall frequency, most words are NOT clustered along the trendline, suggesting there are greater differences in word usage frequencies.

```{r}
ggplot(total6, aes(`technical.fraction`,`writing.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")

@@

@@

##Log-Odds Ratio for Text Comparison Between Text Categories
The log-odds ratios method outlined below finds and displays differences words that are more frequently found in TA comments belonging to one sub-category vs. another. I based this on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](that I found here.) This page has two parts, one for the comparison for code.subject, and a second part for code.structure category.

###Part 1. Log-Odds Ratio Comparisons Between Subject Categories
For simplicity I reduced my analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.

@@

####Pearson Correlations of Log-Odds Word Frequencies
If the words used by TAs in their comments differ between the categories, there should be fairly low correlations overall.

```{r}
#Block below re-organizes the data for comparisons. This pattern sets "2_writing" as the first dataset, then all other data in remaining columns.
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns: word, 2_writing, code.subject, proportion
#Stats need this format to be able to compare 2_writing against other values.
base_data_tokenized_sorted2 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion, `3_technical`:`4_logic`)

@@

#This structure is needed in order to compare word (in Column 1) frequencies for 2. Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes. WHICH subset of X values to use is coded in Column 3.
```

**Pearson correlation** between frequency of words in 2_writing versus 3_technical and 2_writing versus 4_logic sub-categories.

```{r}
cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
         ~ proportion + `2_writing`)
```

```{r}
cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4_logic",], 
         ~ proportion + `2_writing`)
```

```{r}
coroutput23 <-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
         ~ proportion + `2_writing`)
coroutput24 <-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4_logic",], 
         ~ proportion + `2_writing`)
broom::tidy(coroutput23)
broom::tidy(coroutput24)

```{r}
knitr::kable(coroutput23)
```

Comparing sub-groups 3_technical and 4_logic requires recreating the dataframe in slightly different format that excludes the data rows for 2_writing.

```{r}
# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
frequency_writing3 <- filter(base_data, code.subject=="3_technical"|code.subject=="4_logic")
frequency_writing.subcolumns3 <- frequency_writing3 %>% select(1, 22:23)

@@

```{r}
#Block below re-organizes the data for comparisons. This pattern sets "3_technical" as the first dataset, then all other data in remaining columns.
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns: word, 3_technical, code.subject, proportion
#Stats need this format to be able to compare 3_technical against other values.
base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion,`4_logic`)

@@

#This structure is needed in order to compare word (in Column 1) frequencies for 3_technical (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes. WHICH subset of X values to use is coded in Column 3.
```

**Pearson correlation** between the frequency of words in 3_technical versus 4_logic sub-categories.

```{r}
cor.test(data = base_data_tokenized_sorted3[base_data_tokenized_sorted3$code.subject == "4_logic",], 
         ~ proportion + `3_technical`)
```

####Visualizing Word Frequency Differences Between Subject Sub-Categories

```{r}
#Create groups by code.subject, calculates proportional frequency in each group.
#Final step creates 5 columns: code.subject, n, word, total, frequency.
base_data_tokenized_sorted2 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  left_join(base_data_tokenized %>% 
  group_by(code.subject) %>% 
  summarise(total = n())) %>%
  mutate(freq = n/total)
```

```{r}
base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
  select(code.subject, word, freq) %>% 
  spread(code.subject, freq) %>%
  arrange(`2_writing`,`3_technical`)
```

```{r}
ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`3_technical`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
  #ggsave("Freqplot_3vs2.png", width = 10, height = 6)
```

```{r}
#Comparing 2 vs. 4.
base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
  select(code.subject, word, freq) %>% 
  spread(code.subject, freq) %>%
  arrange(`2_writing`,`4_logic`)
```

```{r}
ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`4_logic`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
  #ggsave("Freqplot_4vs2.png", width = 10, height = 6)
```

```{r}
#Create groups by code.subject, calculates proportional frequency in each group.
#Final step creates 5 columns: code.subject, n, word, total, frequency.
base_data_tokenized_sorted3 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  left_join(base_data_tokenized %>% 
  group_by(code.subject) %>% 
  summarise(total = n())) %>%
  mutate(freq = n/total)
```

```{r}
#Set up comparison for 3 vs. 4.
base_data_tokenized_sorted3v4 <- base_data_tokenized_sorted3 %>% 
  select(code.subject, word, freq) %>% 
  spread(code.subject, freq) %>%
  arrange(`2_writing`,`4_logic`)
```

```{r}
ggplot(base_data_tokenized_sorted3v4, aes(`3_technical`,`4_logic`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
  #ggsave("Freqplot_4vs3.png", width = 10, height = 6)
```

####Calculating and Plotting Log-Odds Pairs for Subject Sub-Categories

@@

word_ratios2v4 <- base_data_tokenized %>%
  count(word, code.subject) %>%
  group_by(word) %>%
  filter(sum(n) >= 10) %>%
  ungroup() %>%
  spread(code.subject, n, fill = 0) %>%
  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
  mutate(logratio = log(`2_writing`/`4_logic`)) %>%
  arrange(desc(logratio))
```

```{r}
word_ratios2v3 %>% 
  arrange(abs(logratio))

word_ratios3v4 %>% 
  arrange(abs(logratio))

word_ratios2v4 %>% 
  arrange(abs(logratio))
```

```{r}
word_ratios2v3 %>%
  group_by(logratio < 0) %>%
  top_n(20, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (Writing-top/Technical-bottom)") +
  scale_fill_discrete(name = "", labels = c("2_writing", "3_technical"))
  #ggsave("Logplot_2vs3.png", width = 6, height = 10)

word_ratios3v4 %>%
  group_by(logratio < 0) %>%
  top_n(20, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (Technical-top/Logic-bottom)") +
  scale_fill_discrete(name = "", labels = c("3_technical","4_logic"))
  #ggsave("Logplot_3vs4.png", width = 6, height = 10)

word_ratios2v4 %>%
  group_by(logratio < 0) %>%
  top_n(20, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (Writing-top/Logic-bottom)") +
  scale_fill_discrete(name = "", labels = c("2_writing", "4_logic"))
  #ggsave("Logplot_2vs4.png", width = 6, height = 10)

@@

###Part 2. Log-Odds Ratio Comparisons Between Structure Categories

>>>>

---
title: "Initial Explorations"
output: html_document
---

@@

The first question to answer before making deeper comparisons is **are TAs' comments obviously different between sub-categories?** To that end, I ran three different analyses of word usage in TA comments.

*  Pearson correlations for word use frequencies. This compares words found in all Subject sub-categories
*  Word frequency differences based on log-odds ratios. This highlights terms that are the most different between sub-categories.
*  Unique n-grams mapping. This identifies text strings, phrases, and rare terms that are unique (or nearly unique) to one sub-category.

#Setup
```{r Libraries_1}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

#Read in CSV as a starting dataframe named "base_data". 
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

@@

#Pearson Correlations of Word Frequencies
##Are Shared Words Unequally Distributed?
This analysis takes all comments from base_data, tokenizes and filters stopwords, then counts and sorts fractional frequencies for each word in each category. **NOTE**: this analysis only compares word use for terms present in ALL sub-categories. If a word is MISSING from **any subcategory**, the data row for that word is pruned from the dataset. This eliminates all unique words in favor of words found in all sub-categories.

```{r}
#Tokenize phrases, remove the stop words listed in R's standard reference file.
base_data_tokenized <- base_data %>% 
  unnest_tokens(word,ta.comment)

#ALTERNATIVE METHOD for pulling out stopwords.  
sortedwords.comments.all <- base_data_tokenized %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

#Calculate total # of words
total_words.all <- sortedwords.comments.all %>% 
  summarize(total = sum(n))

#Calculate then append fractional word frequencies.
sortedwords.comments.all.fraction <- sortedwords.comments.all %>% as_tibble() %>% mutate(
  all.fraction = (n*100)/total_words.all$total
  )

#Select SUBSETs of tokenized data based on code.subject: 
base_data_tokenized.basic <- subset(base_data_tokenized, code.subject == "1_basic", select = (1:28))
base_data_tokenized.writing <- subset(base_data_tokenized, code.subject == "2_writing", select = (1:28))
base_data_tokenized.technical <- subset(base_data_tokenized, code.subject == "3_technical", select = (1:28))
base_data_tokenized.logic <- subset(base_data_tokenized, code.subject == "4_logic", select = (1:28))

#For each subset, filter stopwords, count words and total, then calculate frequencies.
sortedwords.basic <- base_data_tokenized.basic %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

@@

#Mutate table to calculate then append fractional word frequency in the text.
sortedwords.basic.fraction <- sortedwords.basic %>% as_tibble() %>% mutate(
  basic.fraction = (n*100)/total_words.basic$total
)
#Repeats for subset 2
sortedwords.writing <- base_data_tokenized.writing %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

total_words.writing <- sortedwords.writing %>% 
  summarize(total = sum(n))

sortedwords.writing.fraction <- sortedwords.writing %>% as_tibble() %>% mutate(
  writing.fraction = (n*100)/total_words.writing$total
)

#Repeats for subset 3
sortedwords.technical <- base_data_tokenized.technical %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

total_words.technical <- sortedwords.technical %>% 
  summarize(total = sum(n))

sortedwords.technical.fraction <- sortedwords.technical %>% as_tibble() %>% mutate(
  technical.fraction = (n*100)/total_words.technical$total
)

#Repeats for subset 4
sortedwords.logic <- base_data_tokenized.logic %>%
  filter(!word %in% stop_words$word,
         !word %in% str_remove_all(stop_words$word, "'"),
         str_detect(word, "[a-z]")) %>%
  count(word, word, sort = TRUE) %>%
  ungroup()

total_words.logic <- sortedwords.logic %>% 
  summarize(total = sum(n))

sortedwords.logic.fraction <- sortedwords.logic %>% as_tibble() %>% mutate(
  logic.fraction = (n*100)/total_words.logic$total
)
```

The next block joins the individual data tables for each subset in a larger file using column named 'word'. Columns get confusing names so they are renamed. 
```{r}
total <- merge(sortedwords.comments.all.fraction,sortedwords.technical.fraction,by="word") 
total2 <- merge(total,sortedwords.writing.fraction,by="word")
total3 <- merge(total2,sortedwords.logic.fraction,by="word")
#rename lines that were possibly confusing
names(total3)[2] <- "all.count"
names(total3)[4] <- "technical.count"
names(total3)[6] <- "writing.count"
names(total3)[8] <- "logic.count"

@@

The following six graphs compare frequencies for ~770 words that appear in ALL three sub-categories: Writing, Technical, and Logic.
```{r}
ggplot(total6, aes(`all.fraction`,`writing.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")

@@

ggplot(total6, aes(`all.fraction`,`logic.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")

ggplot(total6, aes(`technical.fraction`,`writing.fraction`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")

@@

Graphs 1-3 show frequencies in individual groups versus word frequency overall. Generally, frequency of terms in the whole set of comments should be similar to frequency in an individual group (X vs Y positions similar.) However, some terms associated with Logic subgroup (Graph 3) are far from the best-fit line, suggesting they are under-used for comments in that sub-category.

Graphs 4-6 show pair-wise comparisons between sub-categories. Unlike overall frequency, most words are NOT clustered along the best-fit line, suggesting there is still more variation in word use between sub-categories.

@@

#Log-Odds Ratios
##Which Words Are Shared LEAST Between Categories?
Unlike the preceding analysis, log-odds ratio comparison does not eliminate unshared terms. Instead, it identifies which words are MOST likely to be used in one sub-category, but not another. For simplicity I reduced the analysis to just the four most common comment subjects. This has been the most informative preliminary visualization I have found so far.

This code block is based on a method for comparing word usage frequencies between two separate Twitter users [https://www.tidytextmining.com/twitter.html](described here.) If the words used by TAs in their comments differ between sub-categories, there should be fairly low Pearson correlations between word frequencies.

@@

#Block below re-organizes the data for comparisons. This pattern sets "2_writing" as the first dataset, then all other data in remaining columns.
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns: word, 2_writing, code.subject, proportion
#Stats need this format to be able to compare 2_writing against other values.
base_data_tokenized_sorted2 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion, `3_technical`:`4_logic`)

@@

#This structure is needed in order to compare word (in Column 1) frequencies for 2. Writing Quality (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes. WHICH subset of X values to use is coded in Column 3.

#Pearson correlation between frequency of words in 2_writing versus 3_technical and 2_writing versus 4_logic sub-categories.
cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
         ~ proportion + `2_writing`)

cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4_logic",], 
         ~ proportion + `2_writing`)

coroutput23 <-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "3_technical",], 
         ~ proportion + `2_writing`)
coroutput24 <-cor.test(data = base_data_tokenized_sorted2[base_data_tokenized_sorted2$code.subject == "4_logic",], 
         ~ proportion + `2_writing`)

#Comparing sub-groups 3_technical and 4_logic requires recreating the dataframe in a slightly different format that excludes the data rows for 2_writing.

# Isolate table rows to compare. Then reduce larger dataframe to only 2 required columns of data.
frequency_writing3 <- filter(base_data, code.subject=="3_technical"|code.subject=="4_logic")
frequency_writing.subcolumns3 <- frequency_writing3 %>% select(1, 22:23)

@@

#Block below re-organizes the data for comparisons. This pattern sets "3_technical" as the first dataset, then all other data in remaining columns.
#Groups by code.subject, calculated proportional frequency in each group.
#Final step creates 4 columns: word, 3_technical, code.subject, proportion
#Stats need this format to be able to compare 3_technical against other values.
base_data_tokenized_sorted3 <- base_data_tokenized3 %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  mutate(proportion = n / sum(n)) %>% 
  select(-n) %>% 
  spread(code.subject, proportion) %>% 
  gather(code.subject, proportion,`4_logic`)

@@

#This structure is needed in order to compare word (in Column 1) frequencies for 3_technical (Column 2) as Y values against the frequencies of those same words (in Column 4) as X axes. WHICH subset of X values to use is coded in Column 3.

#Pearson correlation** between the frequency of words in 3_technical versus 4_logic sub-categories.
coroutput34 <-cor.test(data = base_data_tokenized_sorted3[base_data_tokenized_sorted3$code.subject == "4_logic",], 
         ~ proportion + `3_technical`)

broom::tidy(coroutput23)
broom::tidy(coroutput24)
broom::tidy(coroutput34)
```

###Visualizing Frequency Differences Between Sub-Categories

```{r}
#Create groups by code.subject, calculates proportional frequency in each group.
#Final step creates 5 columns: code.subject, n, word, total, frequency.
base_data_tokenized_sorted2 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  left_join(base_data_tokenized %>% 
  group_by(code.subject) %>% 
  summarise(total = n())) %>%
  mutate(freq = n/total)

base_data_tokenized_sorted2v3 <- base_data_tokenized_sorted2 %>% 
  select(code.subject, word, freq) %>% 
  spread(code.subject, freq) %>%
  arrange(`2_writing`,`3_technical`)

ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`3_technical`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
  #ggsave("Freqplot_3vs2.png", width = 10, height = 6)

#Comparing 2 vs. 4.
base_data_tokenized_sorted2v4 <- base_data_tokenized_sorted2 %>% 
  select(code.subject, word, freq) %>% 
  spread(code.subject, freq) %>%
  arrange(`2_writing`,`4_logic`)

ggplot(base_data_tokenized_sorted2v3, aes(`2_writing`,`4_logic`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
  #ggsave("Freqplot_4vs2.png", width = 10, height = 6)

#Create groups by code.subject, calculates proportional frequency in each group.
#Final step creates 5 columns: code.subject, n, word, total, frequency.
base_data_tokenized_sorted3 <- base_data_tokenized %>% 
  group_by(code.subject) %>%
  count(code.subject, word, sort=TRUE) %>% 
  left_join(base_data_tokenized %>% 
  group_by(code.subject) %>% 
  summarise(total = n())) %>%
  mutate(freq = n/total)

#Set up comparison for 3 vs. 4.
base_data_tokenized_sorted3v4 <- base_data_tokenized_sorted3 %>% 
  select(code.subject, word, freq) %>% 
  spread(code.subject, freq) %>%
  arrange(`2_writing`,`4_logic`)

ggplot(base_data_tokenized_sorted3v4, aes(`3_technical`,`4_logic`)) +
  geom_jitter(alpha = 0.1, size = 2.5, width = 0.25, height = 0.25) +
  geom_text(aes(label = word), check_overlap = TRUE, vjust = 1.5) +
  scale_x_log10(labels = percent_format()) +
  scale_y_log10(labels = percent_format()) +
  geom_abline(color = "red")
  #ggsave("Freqplot_4vs3.png", width = 10, height = 6)
```

###Calculating and Plotting Log-Odds Pairs for Subject Sub-Categories

@@

word_ratios2v4 <- base_data_tokenized %>%
  count(word, code.subject) %>%
  group_by(word) %>%
  filter(sum(n) >= 10) %>%
  ungroup() %>%
  spread(code.subject, n, fill = 0) %>%
  mutate_if(is.numeric, funs((. + 1) / (sum(.) + 1))) %>%
  mutate(logratio = log(`2_writing`/`4_logic`)) %>%
  arrange(desc(logratio))

word_ratios2v3 %>% 
  arrange(abs(logratio))
word_ratios3v4 %>% 
  arrange(abs(logratio))
word_ratios2v4 %>% 
  arrange(abs(logratio))
```

```{r}
word_ratios2v3 %>%
  group_by(logratio < 0) %>%
  top_n(25, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (Writing-top/Technical-bottom)") +
  scale_fill_discrete(name = "", labels = c("2_writing", "3_technical"))
  #ggsave("Logplot_2vs3.png", width = 6, height = 10)

word_ratios3v4 %>%
  group_by(logratio < 0) %>%
  top_n(25, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (Technical-top/Logic-bottom)") +
  scale_fill_discrete(name = "", labels = c("3_technical","4_logic"))
  #ggsave("Logplot_3vs4.png", width = 6, height = 10)

word_ratios2v4 %>%
  group_by(logratio < 0) %>%
  top_n(25, abs(logratio)) %>%
  ungroup() %>%
  mutate(word = reorder(word, logratio)) %>%
  ggplot(aes(word, logratio, fill = logratio < 0)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  ylab("log odds ratio (Writing-top/Logic-bottom)") +
  scale_fill_discrete(name = "", labels = c("2_writing", "4_logic"))
  #ggsave("Logplot_2vs4.png", width = 6, height = 10)

@@

#N-Grams Mapping 
##Which Terms, Phrases Are Unique to a Sub-Category?

```{}
#Pseudo-code
Import comments
Tokenize the comments as 1-, 2, 3-grams
Add counts within subset, calculate frequency
Split by subcategory
Align 4 subcategories, highest to lowest

####\  
```{r}
# Call in tidyverse, other required libraries.
library(tidyverse)
library(tidytext)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

#Read in CSV as a starting dataframe named "base_data". 
base_data <- read_csv(file='data/coded_full_comments_dataset_Spring18anonVERB.csv')
```

```{r}
#Tokenize phrases as N-grams.

base_data_tokenized2n <- base_data %>% 
  unnest_tokens(ngram, ta.comment, token="ngrams", n=2)

sortedwords.comments.all2n <- base_data_tokenized2n %>%
  count(ngram, sort = TRUE) %>%
  ungroup()

total_words.all2n <- sortedwords.comments.all2n %>% 
  summarize(total = sum(n))

sortedwords.comments.all.fraction2n <- sortedwords.comments.all2n %>% as_tibble() %>% mutate(
  all.fraction = (n*100)/total_words.all2n$total
  )

#Select SUBSETs of tokenized data based on code.subject: 
base_data_tokenized.basic2n <- subset(base_data_tokenized2n, code.subject == "1_basic", select = (1:28))
base_data_tokenized.writing2n <- subset(base_data_tokenized2n, code.subject == "2_writing", select = (1:28))
base_data_tokenized.technical2n <- subset(base_data_tokenized2n, code.subject == "3_technical", select = (1:28))
base_data_tokenized.logic2n <- subset(base_data_tokenized2n, code.subject == "4_logic", select = (1:28))

#For each subset, filter stopwords, count words and total, then calculate frequencies.
sortedwords.basic2n <- base_data_tokenized.basic2n %>%
  count(ngram, sort = TRUE) %>%
  ungroup()
